\def\clim{$C_{lim}$}
\def\hlim{$H_{lim}$}
\def\c2{$C_2$}
\def\im{$I_m$}
\def\km{$K_m$}

% For numbering piecewise equations
\input cases

% Fancy fractions: \displaystyle

\cleardoublepage
\label[methodology]
\chap METHODOLOGY

%IMPORTANT
% Comparing the analytical models of Peter and Cushman can corroborate to my numerical approach in the following form?
% Cushman is MM based and Peter is constant demand. The differences between those analytical models has to be similar as the ones 
% found between my numerical model and the Im constant model. Both analytical models consider steady-state condition for water flow
% so the difference will be due to the uptake rate only.

In this chapter, the model development is described.
First, we show the mathematical equations that represent the transport of water and solutes and their uptake by the root (Section~\ref[model_description]). 
The numerical development with the applied boundary conditions are show in Section~\ref[numerical_implementation]. 
The numerical solutions for the zero \cite[liersolute] and constant uptake 
models -- that were used in the comparison with the proposed model -- are show in Sections \ref[ZU] and \ref[CU] and a comparative analysis with the analytical model of \citeonline[cushman] is made in Section \ref[analytical_models].
Later, we present the description of the simulated scenarios (Section~\ref[scenarios]), a comparative analysis of the linear and non-linear approaches of the proposed model (Section~\ref[lu_nlu_comp]) and its statistical difference (Section \ref[statistics]), and finally, the methodology used for the sensitivity analysis (Section~\ref[sensitivity]).

%In this thesis, we will compare the proposed model with the original one \cite[liersolute], and  with analytical models of constant \cite[willigen1] and concentration dependent (MM) \cite[cushman] uptake, both considering steady-state condition with respect to water flow. 
%A numerical implementation of the constant uptake was developed, therefore, the numerical solutions for the concentration dependent, zero uptake and constant uptake models are shown.

\label[model_description]
\sec General model description

The geometry of the soil-root system considers an uniformly distributed parallel cylindrical root of radius $r_0$ and length $z$. 
To each root, a concentric cylinder of radius $r_m$ and length $z$ can be assigned to represent its extraction volume (Figure \ref[fig_rootzone]).

\medskip
\label[fig_rootzone]
\picw=11cm \cinspic root_zone.pdf
\caption/f {Schematic representation of the spatial distribution of roots in the root zone}
\medskip

The discretization needed for the numerical solution was performed at the single root scale. 
As the extraction properties of the root are considered uniform along its length, and assuming no vertical differences in root density and fluxes, the cylinder can be represented by its cross-section, a circle. 
The area of this circle, representing the extraction region, was subdivided into $n$ circular segments of variable size $\Delta r$~(m), small near the root and increasing with distance, according to the equation \cite[liersolute]:
\label[eq_disc]
$$
\Delta r = \Delta r_{min}+(\Delta r_{max}-\Delta r_{min}) \left( {r-r_0 \over r_m-r_0} \right)^S \eqmark
$$
%
where the subscripts 
in $\Delta r$ 
indicate the minimum and maximum segment sizes defined by the user, and $S$ gives the rate at which the segment size increases. 
The parameter $r_0$ (m) represents the root radius, and $r_m$ (m) is the radius of the root extraction zone, equal to the half-distance between roots, which relates to the root density $R$ (m m$^{-3}$) according to Equation \ref[eq_rm].
This variable size discretization has the advantage to result in smaller segments in regions that need more detail in the calculations (near the root soil interface) due to the greater variation of expected fluxes. 
%It is specially advantageous when dealing with numerical stabilization and convergence problems. 
Figure~\ref[fig_scheme] shows a schematic representation of the discretization as projected by Equation~\ref[eq_disc].

\medskip
\label[fig_scheme]
\picw=9.5cm \cinspic domain.pdf
\caption/f {Schematic representation of the discretized domain considered in the model. $\Delta r$ is the variable segment size, increasing with the distance from the root surface ($r_0$) to the half-distance between roots ($r_m$), and $n$ is the number of segments}
\medskip

A fully implicit numerical treatment was given to the water and solute balance equations \ref[eq_Richards] and \ref[eq_sol].
The Richards equation \ref[eq_Richards] for one-dimensional axisymmetric flow can be written as
\label[eq_complete_richards]
$$
{\partial \theta \over \partial t} = {\partial \theta \over \partial H} {\partial H \over \partial t} = C_w(H) {\partial H \over \partial t} = {1 \over r}{\partial \over \partial r} \left( r K(H,\theta) {\partial H \over \partial r} \right) \eqmark
$$
%
where the total hydraulic head ($H$) is the sum of pressure ($h$) and osmotic ($h_\pi$) heads and $C_w$ (m$^{-1}$) is the differential water capacity (${\partial H \over \partial\theta}$).
Relations between $K$, $\theta$ and $h$ are described by the \citeonline[genuchten80] equation system (Equations \ref[eq_theta] and \ref[eq_K]).
Analogous to \citeonline[vandam_feddes], Equation \ref[eq_complete_richards] can be solved using an implicit scheme of finite differences with the Picard iterative process:
$$
\eqalignno{
\label[eq_disc_water]
{C_w}_i^{j+1,p-1} (H_i^{j+1,p}&-H_i^{j+1,p-1}) + \theta_i^{j+1,p-1}-\theta_i^j = {t^{j+1}-t^j \over r_i \Delta r_i} \times \cr
&\left[ 
r_{i-1/2}K_{i-1/2}^j {H_{i-1}^{j+1,p}-H_i^{j+1,p} \over r_i-r_{i-1}} 
- 
r_{i+1/2}K_{i+1/2}^j {H_{i}^{j+1,p}-H_{i+1}^{j+1,p} \over r_{i+1}-r_{i}} 
\right] &\eqmark
}
$$
%
where $i$ ($1 \leq i \leq n$) refers to the segment number, $j$ is the time step and $p$ the iteration level. 
The Picard's method is used to reduce inaccuracies in the implicit numerical solution for the $h$-based Equation \ref[eq_complete_richards] \cite[celia].

The solution for Equation \ref[eq_disc_water] results in prediction of pressure head in soil as a function of time and distance from the root surface. 
{\localcolor\Red The boundary conditions considered were of flux density entering the root determined by the transpiration rate, for the inner segment; and of zero flux, for the outer segment. {\it QM: Melhor representar essas boundary conditions de forma matematica.}}
The computer algorithm that solves the equation and applied boundary conditions are found on Appendix \ref[ap_water].

The convection-dispersion equation \ref[eq_sol] for one-dimensional axisymmetric flow can be written as
\label[eq_complete_solute]
$$
r {\partial(\theta C) \over \partial t} = -{\partial \over \partial r} \biggl(r q C \biggr) + {\partial  \over \partial r} \left( r D {\partial C \over \partial r} \right). \eqmark
$$
%

Both boundary conditions are of the flux type, according to Equation \ref[eq_boundary_sol].
From the assumed geometry (Figure \ref[fig_scheme]) it follows that the boundary condition at the outer segment corresponds to zero solute flux ($q_s$):
\label[eq_bcrm]
$$
-D(\theta) {\partial C \over \partial r} + q C = q_s = 0\;{\rm ,} \quad t>0,\; r=r_m .\eqmark
$$

The boundary condition for solute transport at the root surface represents the concentration dependent solute uptake, described by the MM equation \ref[eq_MM1], with the following assumptions:
\begitems
\style n
%* Solute uptake by mass flow of water is not controlled by the plant by other means but transpiration, thus the uptake by convection is considered passive;
%* Plant regulates active uptake, thus the uptake by diffusion is plant regulated;
%* Plant demand is equal to the $I_m$ (mol m$^{-2}$ s$^{-1}$) parameter from MM equation;
%* There is a soil solution concentration value (\clim, mol) that the solute flux limits the uptake.
{\localcolor\Red * Solute uptake by mass flow of water is only controlled by the transpiration flow, a convective flow is considered passive;
* Plant regulated active uptake corresponds to diffusion;
* Plant demand is equal to the $I_m$ (mol m$^{-2}$ s$^{-1}$) parameter from the MM equation;
* At a soil solution concentration value \clim (mol), the solute flux limits the uptake. {\it QM: Ainda pode melhorar. Nao sei se estah 100\% claro.}}
\enditems

We assume that the plant demand for solute is constant in time.
The uptake, however, can be higher or lower than the demand, depending on the concentration in the soil solution at root surface (Figure \ref[fig_MM_mod]). 
If the concentration is bellow a certain limiting value (\clim), the uptake is limited by the solute flux, {\it i.e.} solute flux can not attend plant demand even with potential values of active uptake.
Additionally, solute uptake by mass flow of water can be higher than the plant demand in situations of high transpiration rate and/or for high soil water content. 
In these cases, we assume that active uptake is zero and all uptake occurs by the passive process. A concentration \c2 (mol) for this situation is calculated.
When the concentration is between \clim{} and \c2, the uptake is equal to the plant demand as a result of the sum of active and passive contributions to the uptake.
Assumption 1 states that passive uptake is not controlled by any physiological plant mechanisms and, in order to optimize the use of metabolic energy, active uptake is regulated in such way that it works as a complementary mechanism of extraction to achieve plant demand (Assumption 2).
This results in a lower active uptake contribution than that of its potential value.
%--IMPORTANT TO IMPROVE ASSUMPTIONS
%Ions and water move into the plant independently, reflecting entirely different physical-chemical forces; nevertheless, the water moving to the root carries ions in it. A portion of these ions is taken up by the plant by an independent mechanism. The remainder is left behind at the root surface or on the soil particles in the vicinity of the root where they represent a subsequent source of supply. 81.pdf, pg. 98
%This can explain the analysis of diffusion and convection (active and passive) separately.
However, the effect of the solute concentration inside the plant on solute uptake and plant demand is not considered in the model.
Consequently, a scenario for which the demand is reduced due to an excess of solute concentration in the plant is not considered.
This might, in certain situations, lead to an overestimated prediction of uptake.

A piecewise non-linear uptake function that considers these explicit boundary conditions was formulated as:
\label[eq_MM_mod]
\numcases{q_{s0} =}
\label[eq_case1]
{I_m C_0 \over K_m+C_0}+q_0 C_0,&if $C_0 < C_{lim}$ \\
\label[eq_case2]
I_m, 				&if $C_{lim} \leq C \leq C_2$ \\
\label[eq_case3]
q_0 C, 				&if $C > C_2$
\endnumcases
%
with \clim{} determined by the positive root of
\label[eq_clim]
$$
C_{lim} = -{K_m \pm \left( {K_m}^2 + 4{I_m K_m / q_0} \right)^{1/2} \over 2 }, \eqmark
$$
%
and \c2 by
\label[eq_c2]
$$
C_2 = {I_m \over q_0}{\rm .} \eqmark
$$

\medskip
\label[fig_MM_mod]
\picw=13cm \cinspic MM_c.pdf
%\caption/f {Solute uptake piecewise equation \reff_MM_mod after assumptions made over original MM equation \ref[eq_MM1].}
\caption/f {Solute uptake piecewise equation \ref[eq_MM_mod] from MM equation \ref[eq_MM1] with boundary conditions. The bold line represents the actual uptake, thin lines represent active and passive contributions to the actual uptake, and dotted lines represent the plant demand and the potential active uptake}
\medskip

%IMPROVE IT AND EXPLAIN MORE
The non-linear part of the uptake function resides in Equation \ref[eq_case1]. 
As implicit numerical implementations of non-linear functions may result in solutions with stability issues, a linearization of Equation \ref[eq_case1] was made, resulting in:
%Due to the difficulty of implement a numerical algorithm to the non-linear Equation \ref[eq_case1], a linearization was made, resulting in:
\label[eq_MM_linear]
$$
q_{s0} = (\alpha + q_0)\,C_0,\,\,\,\,{\rm if }\, C_0 < C_{lim}  \eqmark
$$
%
where $\alpha$ (m s$^{-1}$) and $q_0$ (m s$^{-1}$) are the active and passive contributions for the solute uptake slope $(\alpha+q_0)$.
%IMPROVE
This linearization is very similar to the one proposed by \citeonline[tinker], but does not consider the solute concentration inside the plant.
The derivation of Equations \ref[eq_clim] to \ref[eq_MM_linear] is shown in Appendix \ref[ap_clim].

Finally, the boundary condition at the inner segment refers to the concentration dependent solute flux at the root surface ($q_{s0}$, mol m$^{-2}$ d$^{-1}$) in agreement to Equation \ref[eq_MM_mod] and \ref[eq_MM_linear] for the non-linear and linear case, respectively.
The uptake of each root equals $-q_{s0}/R$ (mol d$^{-1}$, the negative sign indicating solute depletion), thus, the condition at root surface is described by:
\label[eq_inner_bound]
$$
-D(\theta) {\partial C \over \partial r} + q_0 C_0 = -{q_{s0} \over 2\pi r_0 R z}, \eqmark
%= {I_m C_0 \over K_m+C_0}+q_0 C_0\;{\rm .} \quad t>0,\; r=r_0 \eqmark
$$

{\localcolor\Red
The initial condition corresponds to constant solute concentration ($C_{ini}$) in all segments:
$$
C = C_{ini}\;, \quad t=0,\; r=r_i,\; 1 \leq i \leq n. \eqmark
$$
%
{\it QM:Nao seria melhor colocar isso antes do tudo (comecar com isso)? Ficou um pouco perdido aqui.}
}

\label[numerical_implementation]
\sec Numerical implementations of the convection-dispersion equation

In this item, the numerical discretization and the applied boundary conditions are presented, together with the algorithm to find the solution of Equation \ref[eq_complete_solute] for both linear and non-linear boundary conditions at the root surface.
The mass balance equations are considered to correspond to transient conditions with respect to water and solute.

As mentioned in the introduction of this chapter, besides the development of the proposed model (concentration dependent solute uptake), using the same discretization detailed in the previous section a numerical solution for a constant uptake model based on \citeonline[willigen1] was developed. 
Additionally, we present the numerical solution for the zero uptake model \cite[liersolute], in which some modifications are proposed. 
In this text, zero uptake model will be abbreviated as ZU, constant uptake model as CU, linear concentration dependent model as LU and non-linear concentration dependent model as NLU. 

%MODIFY IT. Copyied from quirijn
In the numerical solutions, the combined water and solute movement is simulated iteratively. In a first step, the water movement towards the root is simulated, assuming salt concentrations from the previous time step. In a second step, the salt contents per segment are updated and new values for the osmotic head in all segments are calculated. The first step is then repeated with updated values for the osmotic heads. This process is repeated until the pressure head values and osmotic head values between iterations converge. 
Flowcharts containing the algorithm structure are shown in the Appendix \ref[ap_flowchart].

\secc Concentration dependent uptake model (proposed model)

The implicit numerical discretization of Equation \ref[eq_complete_solute] yields:
$$
\eqalignno{
\label[eq_disc_sol]
\theta^{j+1}_i &C^{j+1}_i - \theta^j_i C^j_i = {\Delta t \over 2 r_i \Delta r_i} \times \cr
& \left\{ 
{r_{i-1/2} \over r_i-r_{i-1}} \left[ q_{i-1/2}(C^{j+1}_{i-1}\Delta r_i + C^{j+1}_{i}\Delta r_{i-1}) - 2D^{j+1}_{i-1/2}(C^{j+1}_i-C^{j+1}_{i-1}) \right]\right. - & \eqmark \cr
&\left. {r_{i+1/2} \over r_{i+1}-r_{i}} \left[ q_{i+1/2}(C^{j+1}_{i}\Delta r_{i+1} + C^{j+1}_{i+1}\Delta r_{i}) - 2D^{j+1}_{i+1/2}(C^{j+1}_{i+1}-C^{j+1}_{i}) \right] 
\right\}
}
$$

Applying equation \ref[eq_disc_sol] to each segment, the concentrations for the next time step $C_i^{j+1}$ (mol m$^{-3}$) are obtained by solving the following tridiagonal matrix:
\label[matrix]
$$
\left [ \matrix{
 b_1 & c_1 & & & &  \cr
 a_2 & b_2 & c_2 & & &  \cr
  & a_3 & b_3 & c_3 & &  \cr
  &  & \ddots & \ddots & \ddots &  \cr
  &  & & a_{n-1} & b_{n-1} & c_{n-1}  \cr
  &  & & & a_{n} & b_{n}  \cr}  \right]
\left [ \matrix{
 C^{j+1}_1 \cr
 C^{j+1}_2 \cr
 C^{j+1}_3 \cr
 \vdots \cr
 C^{j+1}_{n-1} \cr
 C^{j+1}_n \cr} \right ]
=
\left [ \matrix{
 f_1 \cr
 f_2 \cr
 f_3 \cr
 \vdots \cr
 f_{n-1} \cr
 f_n \cr} \right ]
 \eqmark
$$
%
with $f_i \hbox{(mol m$^{-2}$)}$ defined
%, unless specified otherwise, 
as
$$
f_i = r_i \theta^j_ i C^j_i \eqmark
$$
%
and $a_i\, \hbox{(m)}$, $b_i\, \hbox{(m)}$ and $c_i\, \hbox{(m)}$ defined for the respective segments 
%and model type 
as described in the following.

\seccc The intermediate nodes ($i = 2\hbox{ to }i = n-1$)

Rearrangement of Equation \ref[eq_disc_sol] to \ref[matrix] results in the coefficients:
$$
\eqalignno{
\label[eq_int_nodes1]
a_i &= -{r_{i-1/2} (2D^{j+1}_{i-1/2} + q_{i-1/2} \Delta r_i)\Delta t \over 2(r_{i}-r_{i-1})\Delta r_i} & \eqmark \cr
b_i &= r_i \theta^{j+1}_i + {\Delta t \over 2 \Delta r_i} 
\left[
{r_{i-1/2} \over (r_i-r_{i-1})} (2D^{j+1}_{i-1/2} - q_{i-1/2} \Delta r_{i-1}) + \right. \cr
&\left.{r_{i+1/2} \over (r_{i+1}-r_{i})} (2D^{j+1}_{i+1/2} + q_{i+1/2} \Delta r_{i+1}) 
\right] & \eqmark \cr
\label[eq_int_nodes3]
c_i &= -{r_{i+1/2} \Delta t \over 2 \Delta r_i (r_{i+1}-r_i)} (2D^{j+1}_{i+1/2} - q_{i+1/2} \Delta r_i) & \eqmark \cr
}
$$

\seccc The outer boundary ($i=n$)

Applying boundary condition of zero solute flux, the third and fourth terms from the right hand side of Equation \ref[eq_disc_sol] are equal to zero. Thus, the solute balance for this segment is written as:
$$
\eqalignno{
\label[outer]
\theta^{j+1}_n &C^{j+1}_n - \theta^j_n C^j_n = {\Delta t \over 2 r_n \Delta r_n} \times \cr
& \left\{ 
{r_{n-1/2} \over r_n-r_{n-1}} \left[ q_{n-1/2}(C^{j+1}_{n-1}\Delta r_n + C^{j+1}_{n}\Delta r_{n-1}) -  2D^{j+1}_{n-1/2}(C^{j+1}_n-C^{j+1}_{n-1}) \right]
\right\} & \eqmark
}
$$

Rearrangement of Equation \ref[outer] to \ref[matrix] results in the coefficients:
$$
\eqalignno{
a_n &= -{r_{n-1/2} (2D^{j+1}_{n-1/2} + q_{n-1/2} \Delta r_n)\Delta t \over 2(r_{n}-r_{n-1})\Delta r_n} & \eqmark \cr
b_n &= r_n \theta^{j+1}_n + {\Delta t \over 2 \Delta r_n} 
\left[
{r_{n-1/2} \over (r_n-r_{n-1})} (2D^{j+1}_{n-1/2} + q_{n-1/2} \Delta r_{n-1})
\right] & \eqmark
}
$$

\seccc The inner boundary ($i=1$)

\begitems
\style a
* {\bf Linear concentration dependent model (LU)}

Applying boundary conditions of linear concentration dependent solute flux, the first and second terms of the right-hand side of Equation \ref[eq_disc_sol] are equal to 
$\displaystyle{-{\alpha + q_0 \over 2\pi r_o R z}}C^{j+1}_1 \Delta r_1$ while $C < C_{lim}$ and $C > C_2$:
$$
\eqalignno{
\label[LU]
\theta^{j+1}_1 &C^{j+1}_1 - \theta^j_1 C^j_1 = {\Delta t \over 2 r_1 \Delta r_1} \times \cr
& \left\{  
{r_{1-1/2} \over r_{1}-r_{0}} \left[-{\alpha + q_0 \over 2\pi r_0 R z}\right] C^{j+1}_1 \Delta r_1 \right. - & \eqmark \cr
& \left. { r_{1+1/2} \over r_{2}-r_{1}} \left[ q_{1+1/2}(C^{j+1}_{1}\Delta r_{2} + C^{j+1}_{2}\Delta r_{1}) - 2D^{j+1}_{1+1/2}(C^{j+1}_{2}-C^{j+1}_{1}) \right] 
\right\} &
}
$$
%
When $C=0$ the solute flux is set to zero and the equation is equal to Equation \ref[lier] (zero uptake). While $C_{lim}~\leq~C~\leq~C_2$, the solute flux density is constant and the equation is equal to Equation \ref[will] (constant uptake).

The introduction of Equation \ref[LU] in the tridiagonal matrix \ref[matrix] results in the following coefficients:
$$
\eqalignno{
b_1 &= r_1 \theta^{j+1}_1 + {\Delta t \over 2 \Delta r_1} \left[ {r_{1+1/2} \over (r_2-r_1)} (2D^{j+1}_{1+1/2} + q_{i+1/2} \Delta r_2) + {r_{1-1/2} \over (r_{1}-r_{0})}{\left(\alpha + q_0 \right) \over 2\pi r_0 R z} \Delta r_1\right] & \eqmark \cr 
c_1 &= -{r_{1+1/2} \Delta t \over 2 \Delta r_1 (r_2-r_1)} (2D^{j+1}_{1+1/2} - q_{1+1/2} \Delta r_1) & \eqmark
}
$$
% bs(1) = dt/(2.*dr(1)) * ((Imm/c_lim*dr(1)+q(0)*dr(1))/(2.*PI*R*z*r(0)) + rsup(1)/(r(2)-r(1)) * (2*D(1)+q(1)*dr(2)) )+ r(1)*theta(1)

* {\bf Non-linear concentration dependent model (NLU)}

Applying boundary conditions of non-linear concentration dependent solute flux, the first and second term of the right-hand side of Equation \ref[eq_disc_sol] become 
$\displaystyle{-\left({I_m \over 2 \pi r_0 R z(K_m +C^{j+1}_1)} + q_0\right)}C^{j+1}_1 \Delta r_1$ for $C < C_{lim}$ and $C > C_2$:
$$
\eqalignno{
\label[NLU]
\theta^{j+1}_1 &C^{j+1}_1 - \theta^j_1 C^j_1 = {\Delta t \over 2 r_1 \Delta r_1} \times \cr
& \left\{  
{r_{1-1/2} \over r_{1}-r_{0}} \left[-\left({I_m \over 2 \pi r_0 R z (K_m +C^{j+1}_1)} + q_0\right)\right] C^{j+1}_1 \Delta r_1 \right. - & \eqmark \cr
& \left. { r_{1+1/2} \over r_{2}-r_{1}} \left[ q_{1+1/2}(C^{j+1}_{1}\Delta r_{2} + C^{j+1}_{2}\Delta r_{1}) - 2D^{j+1}_{1+1/2}(C^{j+1}_{2}-C^{j+1}_{1}) \right] 
\right\} & 
}
$$

Rearrangement of Equation \ref[NLU] to \ref[matrix] results in the following coefficients:
$$
\eqalignno{
\label[b1]
b_1 = r_1 \theta^{j+1}_1 + {\Delta t \over 2 \Delta r_1} &\left[ {r_{1+1/2} \over (r_2-r_1)} (2D^{j+1}_{1+1/2} + q_{i+1/2} \Delta r_2) + \right. \cr
& \left. {r_{1-1/2} \over r_{1}-r_{0}}\left({I_m \over 2 \pi r_0 R z (K_m +C^{j+1}_1)} + q_0 \right) \Delta r_1\right] & \eqmark \cr
c_1 = -{r_{1+1/2} \Delta t \over 2 \Delta r_1 (r_2-r_1)} &(2D^{j+1}_{1+1/2} - q_{1+1/2} \Delta r_1) & \eqmark
}
$$

\enditems

The value of $C^{j+1}_1$ in Equation \ref[b1] is found using the iterative Newton-Raphson method. 
%Note that since this is a one-dimensional microscopic model, it is assumed that the root has the same characteristcs in all vertical soil profile (along its vertical axis), thus, water and solute transport from soil towards the roots and uptakes are occurring at the same rate in the vertical profile. It is possible to couple this model in another one that has discretized soil layers. For a 2D model (depth and radial distance), the solutions presented here are applied in each layer independently. For a 1D model (only depth), an average of water and solute content through the horizontal profile has to be determined.

\label[ZU]
\secc Zero uptake model (ZU)

The zero uptake solution is the original model presented by \citeonline[liersolute]. The numerical discretization is according to Equation \ref[eq_disc_sol] and the intermediate nodes are analogous to Equations \ref[eq_int_nodes1] to \ref[eq_int_nodes3]. 
The only difference is the boundary at root surface (Equation~\ref[eq_inner_bound]), which is of zero solute flux:
$$
q_{s0}=0 \eqmark
$$

\seccc The inner boundary ($i=1$)

Applying boundary condition of zero solute flux, the first and second term of the right-hand side of Equation \ref[eq_disc_sol] are equal to zero:a
$$
\eqalignno{
\label[lier]
\theta^{j+1}_1 &C^{j+1}_1 - \theta^j_1 C^j_1 = {\Delta t \over 2 r_1 \Delta r_1} \times \cr
& \left\{ 
{r_{1+1/2} \over r_{2}-r_{1}} \left[ -q_{1+1/2}(C^{j+1}_{1}\Delta r_{2} + C^{j+1}_{2}\Delta r_{1}) + 2D^{j+1}_{1+1/2}(C^{j+1}_{2}-C^{j+1}_{1}) \right] 
\right\} & \eqmark
}
$$

Introduction of Equation \ref[lier] in the tridiagonal matrix \ref[matrix] results in the following coefficients:
$$
\eqalignno{
b_1 &= r_1 \theta^{j+1}_1 + {\Delta t \over 2 \Delta r_1} \left[ {r_{1+1/2} \over (r_2-r_1)} (2D^{j+1}_{1+1/2} + q_{1+1/2} \Delta r_2) \right] & \eqmark \cr
c_1 &= -{r_{1+1/2} \Delta t \over 2 \Delta r_1 (r_2-r_1)} (2D^{j+1}_{1+1/2} - q_{1+1/2} \Delta r_1) & \eqmark \cr
}
$$

\label[CU]
\secc Constant uptake model (CU)

The constant uptake solution is based on the model proposed by \citeonline[willigen1]. 
The numerical discretization takes into consideration Equation \ref[eq_disc_sol], whereas the intermediate nodes are analogous to Equations \ref[eq_int_nodes1] to \ref[eq_int_nodes3]. 
The boundary condition at the root surface (Equation~\ref[eq_inner_bound]) corresponds to constant solute flux:
$$
q_{s0}=-{I_m \over 2\pi r_0 R z}. \eqmark
$$

\seccc The inner boundary ($i=1$)

Applying boundary conditions of constant solute flux, the first and second term of the right-hand side of Equation \ref[eq_disc_sol] become 
$\displaystyle{-{I_m \over 2 \pi r_0 R z}}\Delta r_1$ for $C > 0$:
$$
\eqalignno{
\label[will]
\theta^{j+1}_1 &C^{j+1}_1 - \theta^j_1 C^j_1 = {\Delta t \over 2 r_1 \Delta r_1} \times \cr
& \left\{  
{r_{1-1/2} \over r_{1}-r_{0}} \left(-{I_m \over 2 \pi r_0 R z}\right) \Delta r_1 - \right. & \eqmark \cr 
& \left. { r_{1+1/2} \over r_{2}-r_{1}} \left[ q_{1+1/2}(C^{j+1}_{1}\Delta r_{2} + C^{j+1}_{2}\Delta r_{1}) - 2D^{j+1}_{1+1/2}(C^{j+1}_{2}-C^{j+1}_{1}) \right] 
\right\} & 
}
$$
%
When $C = 0$ the solute flux is set to zero and equation \ref[will] reduces to Equation \ref[lier].

Introduction of Equation \ref[will] in the tridiagonal matrix \ref[matrix] results in the following coefficients:
$$
\eqalignno{
b_1 &= r_1 \theta^{j+1}_1 + {\Delta t \over 2 \Delta r_1} \left[ {r_{1+1/2} \over (r_2-r_1)} (2D^{j+1}_{1+1/2} + q_{1+1/2} \Delta r_2) \right] & \eqmark \cr
c_1 &= -{r_{1+1/2} \Delta t \over 2 \Delta r_1 (r_2-r_1)} (2D^{j+1}_{1+1/2} - q_{1+1/2} \Delta r_1) & \eqmark \cr
f_1 &= r_1 \theta^j_1 C^j_1 - {r_{1-1/2} \over r_{1}-r_{0}} I_m {\Delta t \over 4 \pi r_0 R z} & \eqmark
}
$$


\label[analytical_models]
\sec Comparison with an analytical model

A computational implementation of the analytical model of \citeonline[cushman] was performed to confront it with the non-linear numerical solution (NLU). 
This particular analytical model also uses the MM equation as boundary condition for solute uptake at the root surface.
The equation by \citeonline[cushman] is:

\label[eq_cush]
$$
\Theta(\mu,\eta) = \sum_{n=0}^\infty A_n \mu^v \beta(\mu,\tau,\alpha_n) \exp(-\alpha_n^2\eta)+1 \eqmark
$$
where
$$
\eqalignno{
\beta_v(\mu,\tau,\alpha_n) &= \left[\alpha_n aY_{v-1}(a\alpha_n)-2v(1-\tau)Y_v(\alpha_n)\right] J_v(\alpha_n\mu) & \eqmark\cr
                         &- \left[\alpha_n aJ_{v-1}(a\alpha_n)-2v(1-\tau)J_v(\alpha_n)\right] Y_v(\alpha_n\mu),
}
$$
the $\alpha_n$ are roots of
$$
\eqalignno{
&[\alpha aJ_{v-1}(\alpha a)-2vJ_v(\alpha a)] [\alpha Y_{v-1}(\alpha)-2v(1-\tau)Y_v(\alpha)] + & \eqmark\cr
&[2vY_{v}(\alpha a)-\alpha aY_{v-1}(\alpha a)] [\alpha J_{v-1}(\alpha)-2v(1-\tau)J_v(\alpha)] = 0,
}
$$
%
and the $A_n$ are given by
$$
A_n=-{\displaystyle\int_1^a \mu^{1-v}\beta_v(\mu,\tau,\alpha_n)d\mu \over \displaystyle\int_1^a \mu{\beta_v}^2(\mu,\tau,\alpha_n)d\mu}. \eqmark
$$

This analytical solution corresponds to a linear boundary condition. 
The simplification was based on the assumption that the solute concentration is initially low, making the non-linear MM equation to become linear:
\label[eq_cond_linear]
$$
{kC \over 1+kC J_{max}} \approx kC \eqmark
$$
and the boundary condition to become
\label[eq_bc_cush]
$$
D{\partial C \over \partial r}+q_0 C = kC. \eqmark
$$

The non-dimensional variables 
{\localcolor\Red $\eta$, $\mu$, $\tau$, $\Theta$, $v$ and $a$ account for time, distance, root ability to absorb solute by convective flux, solute concentration in soil solution, ratio of convective to diffusive flux and outer influence of the root, respectively.
Parameter $k$, inside the equation $\tau=k/q_0$, is the root absorbing power and equal to $K_m$ in the nomenclature of this thesis.
{\it QM: Adicionar unidades ou dimensoes ajudaria nessa secao.}}


%Here we present the analytical models used in the comparisons to the proposed model. An computational implementation were made to each model.

%{\localcolor\Red MAYBE IT IS NOT NECESSARY TO SHOWN THE ANALYTICAL EQUATIONS, JUST CITE THEM}

\label[scenarios]
\sec Simulation scenarios

The simulations were performed using the hydraulic parameters from the Dutch Staring series \cite[wosten] for three different soils types, as listed in Table~\ref[soils]. 
The general system parameters for the different scenarios are listed in Table \ref[general_param] and values for the Michaelis-Menten (MM) parameters in Table~\ref[MMparam]. 
Values of root length density, initial solute concentration, relative transpiration, soil type, and ion species were chosen at several values, composing eight distinct scenarios as listed in Table \ref[tab_scenarios]. 
Scenario 1 was considered as default, the other scenarios derive from scenario 1 by changing only one input parameter. 
In this way, the effect of variation in soil hydraulic properties is exemplified by scenarios 1, 6 and 7; root length density by scenarios 1, 4 and 5; initial solute concentration by scenarios 1 and 3; and potential transpiration by scenarios 1 and 2.
%The chosen MM parameters were for K$^+$ solute.

\input tables/soils

\input tables/general_param

\input tables/MMparam

\input tables/scenarios

Values of $\Delta r_{min}$, $\Delta r_{max}$ and $S$ in Equation \ref[eq_disc] were 10$^{-5}$ m, 5~10$^{-4}$ m and 0.5, resulting in 22, 68 and 213 segments for the high, medium and low root density simulations, respectively.
To guarantee complete convergence for the non-linear model, a time step of 0.01 s was used when $C_0 < C_{lim}$.

\label[lu_nlu_comp]
\sec Analysis of linear and non-linear approaches

To analyze the differences between the two proposed models (linear and non-linear solutions), the relative differences in the predicted concentrations ($di\!f\!\!f_{C}$) and accumulated uptake ($di\!f\!\!f_{Ac}$), for both models, were calculated as follows:
$$
\eqalignno{
\label[error_C]
di\!f\!\!f_{C}  &= {\sum_{x=1}^{x_{end}} C\!L_x - C\!N\!L_x \over \sum_{x=1}^{x_{end}} C\!L_x} & \eqmark \cr
\label[error_acc] 
di\!f\!\!f_{Ac} &= {\sum_{t=1}^{t_{end}} AcL_t - AcN\!L_t \over \sum_{t=1}^{t_{end}} AcL_t} & \eqmark \cr
}
$$
%
where $C\!L_x$ and $C\!N\!L_x$, are the solute concentration in soil water, and $AcL_t$ and $AcN\!L_t$ the accumulated uptake, for LU and NLU, respectively.  
$x$ can be the time ($t$) or the distance from the axial center ($r$). 
The relative difference between three outputs was computed: two relative to time -- concentration at root surface $C_0(t)$ and accumulated solute uptake $Ac(t)$ -- and one relative to radial distance -- concentration $C(r)$.

%The analysis of the results was made in order to choose one out of the two models in further simulations. 
NLU solution uses the non-linear MM equation \ref[eq_MM_mod] and, due to an additional iterative process in the numerical implementation, more time is needed to compute the results when compared with the linear solution LU. 
It is also susceptible to numerical stability issues, depending on selected time and space steps. 
On the other hand, LU is a simplified version of the MM equation in a way that the solute uptake rate for $C_0 < C_{lim}$ is always smaller than that of the original non-linear equation (details in Appendix \ref[ap_l_nl_fluxes]). 
It has no stability problems and needs less computational time because it is less sensitive to space and time steps. 
In a first analysis, the objective was to check if the difference in the results generated by the linearization of the MM equation is sufficiently large to be properly analyzed.  
To do so, four different scenarios were chosen (scenarios 1 to 4 as listed in Table \ref[tab_scenarios]).


%\begitems
%%\style i
%* Scenario 1: Medium root length density, High concentration and High potential transpiration, Loam (default)
%* Scenario 2: Medium root length density, High concentration and Low potential transpiration, Loam
%* Scenario 3: Medium root length density, Low concentration and High potential transpiration, Loam
%* Scenario 4: High root length density, High concentration and High potential transpiration, Loam
%* Scenario 5: Low root length density, High concentration and High potential transpiration, Loam
%* Scenario 6: default, Sand
%* Scenario 7: default, Clay
%* Scenario 8: default, Loam, Other solute (different $I_m$ and $K_m$)
%\enditems

%{\localcolor\Red
%However, further reductions of the segment size might be needed in case of stabilization problems in the solute transport equation.
%TELL EXPLICITLY IN WHICH CASES THE SEGMENT SIZE WAS REDUCED AND THEIR RESPECTIVE VALUES
%}

\label[statistics]
\sec Statistical difference

The Mann--Whitney U test \cite[mcknight] was performed with three datasets for the two models and for each scenario:
(1) concentration over time -- $C_0(t)$;
(2) accumulated uptake over time -- $Ac(t)$;
(3) concentration over radial distance -- $C(r)$). 
This test was chosen because the frequency distributions for both datasets are not normal, the pairs (datasets for LU and NLU) are distinct and do not affect each other. 
This test can distinguish whether each pair is identical or not without assuming them to follow the normal distribution (nonparametric test). 
The null hypothesis ($H_0$) is that both groups 
%(model output data) 
do not differ from each other and the alternate hypothesis ($H_1$) is that groups are different.

The interpretation of test results allows to decide in which cases a particular model should be used (in case they are different) or if both models are equally useful (in case they present no statistical difference).

%\label[model_comp]
%\sec Model comparisons
%
%The scenario of this simulation is of loam soil, medium root length density, high potential transpiration and high initial concentration (Table \ref[general_param]). 
%We compare all model types (no solute uptake -- NU; constant -- CU; linear -- LU and non-linear -- NLU concentration dependent uptake rates). 
%All simulations were made until the value of relative transpiration was equal or less than 0.001. 
%The time step is dynamic (depends on the number of iterations of the previous time step for water and solute equations) and was set to vary between 0.1 and 2 seconds. 
%The simulation for NU ended within near 3 days; for CU, LU and NLU, about 5 days.
%
%%CONSIDERATIONS:
%%•	It is not considered root growth
%%•	It is not considered movement of solute inside the plant
%%•	There is no maximum uptake. The Im value is considered to be the plant demand. The maximum flux is dependent on water flux times concentration in soil solution at root surface
%%•	The model does not act like as an infinite sink (all solute that arrives the root is extracted)
%%•	The model does not include extraction by root hairs. In some cases (like extraction of P) it can contribute 60% to 90% of the uptake
%%•	It does not consider complementary ion effects
%%•	It does not consider soil buff power
%%•	See 21.pdf for limitations in measurement of MM parameters
%
\label[sensitivity]
\sec Sensitivity analysis

The relative partial sensitivity ($\eta$) \cite[quirijn_sens] of model predictions 
%for tend, C0(tend), av. h, av. hos, h(tend) and hos(tend) 
($Y$) as a function of the respective parameter value ($P$) was calculated as
\label[eq_sens]
$$
\eta = {dY/Y \over dP/P}\; \eqmark
$$
%
where $P$ is the default value of the parameter, $dP$ is the in(de)crement applied to $P$, $Y$ is the output of a selected predicted variable and $dY$ is the variation over $Y$ when applied the new parameter value $P \pm dP$.

To determine the sensitivity of the model to an input parameter, the magnitude of its derivative in respect of the model result is calculated. 
If this derivative is close to zero, the model has a low sensitivity to the respective parameter. 
The higher the derivative, the higher is the sensitivity and, therefore, the higher is the precision required when determining that parameter. 
By making a relative analysis like in Equation \ref[eq_sens], the sensitivity for distinct parameters can be compared. 

To determine the sensitivity, a $dP/P$ of 0.01 (1\%) was used for the following selected parameters: 
a) MM parameters $I_m$, $K_m$; and  
b) soil hydraulic parameters $\alpha$, $n$, $\lambda$, $K_s$, $\theta_r$ and $\theta_s$.
The analyzed predicted variables ($Y$) were: end time of simulation $t_{end}$, osmotic head at end of simulation $h_{\pi}$, pressure head at end of simulation $h$, average osmotic head in the soil profile at end of simulation $\overline{h_{\pi}}$, average pressure head in soil profile at end of simulation $\overline{h}$ and accumulated uptake at end of simulation $Ac$.
%
%{\localcolor\Red ::DRAFT:: IMPROVE IT!}
%
%Since the sensitivity analisys is just to analise the derivative of the function prediction X parameter values, we are interested in the slope of this function. So, an infinitesimal variation of $P$ ($dP$) is straight-forward manner to do it.
%
%A variation of 1\% (dP/P=0.01) of $I_m$ and $K_m$ where made from the default values of K$^+$ ion and the relative partial sensitivity was calculated for: end time of simulation ($t_{end}$, s); pressure ($h_{end}$, m) and osmotic ($h_{\pi_{end}}$, m) heads at root surface in the end time of simulation.
%
%
%\medskip
%\label[fig_saltsoilconc]
%\picw=11cm \cinspic productivityXsalinity.pdf
%\caption/f {Biomass production of four diverse and important plant species in a range of salinities. Wheat is one of the more salt-tolerant crops, and rice is one of the more salt-sensitive crops. Source: \cite[munns]}
%\medskip
%
%Biomass production of four diverse and important plant species in a range of salinities. Wheat is one of the more salt-tolerant crops, and rice is one of the more salt-sensitive crops. 
%
