% The documentation of the usage of TeXSALQ -- the template for
% typessetting thesis by plain\TeX at ESALQ - USP
% ---------------------------------------------------------------------
% Luciano R Silveira  Jun.-Jul. 2014
% lrsilveira@gmail.com
%
% Based on ctustyle-doc.tex by Petr Olsak

\input macros/texsalq   % The template is included here.

\worktype [D/EN] % Type: M = master, D = Ph.D., O = other
                 % / the language: PT = Portuguese, EN = English

\title    {Mechanistic numerical modeling of solute uptake by plant roots}
\author     {Andre Herman Freire Bezerra}
\authorinfo {Bachelor in Agronomy}
\city       {Piracicaba}
\date       {2015}
\supervisor {Prof. Dr. \bf\uppercase{Quirijn de Jong van Lier}}  % One or more supervisors
\workname   {} % Used only if \worktype [O/*] (Other)

            % Title / Subtitle in minor language:
\titlePT      {Modelagem numérica de extração de solutos pelas raízes}
\titarea    {Ciências}
\concarea   {Engenharia de Sistemas Agrícolas}
\concareaEN {Agricultural Systems Engineering}
\pagetwo    {}  % The text printed on the page 2 at the bottom.
\dedication {      %  Optional. Use main language here
     To the past, \par\vskip 0.3cm
    to the present, and \par\vskip 0.3cm
    to the future \par\vskip 1.3cm
    %With love%, \bfit DEDICO\par
}
\thanks     {      %  Optional. Use main language here
%Agradeço aos meus amigos e colegas de pós-graduação...
%
%Ao Conselho Nacional de Densenvolvimento Científico e Tecnológico (CNPq) pela bolsa de doutorado e à Coordenação de Aperfeiçoamento de Pessoal de Nível Superior (CAPES) pela bolsa de estágio {\it sandwich}.
%
%Ao meu orientador Quirijn de Jong van Lier por...
%
%I also would like to thank my coleagues at the Wageningen University for all the support that, even without knowing, they gave me and specially to my advisors Sjoerd van der Zee (official), Jos van Dam and Peter de Willigen to every contribution for this work. 
%Wihtout your support, this thesis would be still a project.
    }
\epigraph   {       %  Optional. Use main language here
   {\sl
Some catch phrase that everyone expect to read.
   }

}
\abstractEN {%
A modification in an existing water uptake and solute transport numerical model was implemented in order to allow the model to simulate solute uptake by the roots.
The convection-dispersion equation (CDE) was solved numerically, using a complete implicit scheme, considering a transient state for water and solute fluxes and a soil solute concentration dependent boundary for the uptake at the root surface, based on the Michaelis-Menten (MM) equation.
Additionally, a linear approximation was developed for the MM equation such that the CDE has a linear and a non-linear solution.
A radial geometry was assumed, considering a single root with its surface acting as the uptake boundary and the outer boundary being the half distance between neighboring roots, a function of root density.
The proposed solute transport model includes active and passive solute uptake and predicts solute concentration as a function of time and distance from the root surface.
It also estimates the relative transpiration of the plant, on its turn directly affecting water and solute uptake and related to water and osmotic stress status of the plant.
Performed simulations show that the linear and non-linear solutions result in significantly different solute uptake predictions when the soil solute concentration is below a limiting value ($C_{lim}$).
This reduction in uptake at low concentrations may result in a further reduction in the relative transpiration.
%, {\localcolor\Red shortening the time at which the plant ceases the uptake. {\it QM: This is not clear. Change words}}
The contributions of active and passive uptake vary with parameters related to the ion species, the plant, the atmosphere and the soil hydraulic properties. 
%{\localcolor\Red As the model has the ability to partition active and passive uptake, it can be improved to differentiate osmotic and ionic stress by including a solute concentration inside the plant variable and compare it with a threshold value of plant concentration, which is regulated by plant age and species, for example.}
The model showed a good agreement with an analytical model that uses a linear concentration dependent equation as boundary condition for uptake at the root surface.
The advantage of the numerical model is it allows simulation of transient solute and water uptake and, therefore, can be used in a wider range of situations.
Simulation with different scenarios and comparison with experimental results are needed to verify model performance and possibly suggest improvements.
}
\abstractPT {%
Uma modificação em um modelo existente de extração de água e transporte de solutos foi realizada com o objetivo de incluir nele a possibilidade de simular a extração de soluto pelas raízes.
Uma solução numérica para a equação de convecção-dispersão (ECD), que utiliza um esquema de resolução completamente implícito, foi elaborada e considera o fluxo transiente de água e solutos com uma condição de contorno à superfície da raiz de extração de soluto dependente de sua concentração no solo, baseada na equação de Michaelis-Menten (MM).
Uma aproximação linear para a equação de MM foi implementada de tal forma que a ECD tem uma solução linear e outra não-linear.
O modelo considera uma raiz singular com geometria radial sendo sua superfície a condição de contorno (limite) de extração e sendo o limite extremo a meia-distância entre raízes vizinhas, função da densidade radicular.
O modelo de transporte de soluto proposto inclui extração de soluto ativa e passiva e prediz a concentração de soluto como uma função do tempo e da distância à superfície da raiz,
além de estimar a transpiração relativa da planta, que por sua vez afeta a extração de água e solutos e é relacionado com a condição de estresse da planta.
Simulações mostram que as soluções linear e não-linear resultam em predições de extração de solutos significativamente diferentes quando a concentração de solutos no solo está abaixo de um valor limitante ($C_{lim}$).
A redução da extração em baixas concentrações pode resultar em uma redução adicional na transpiração relativa.
As contribuições ativa e passiva da extração de solutos variam com parâmetros relacionados à espécie de íon, à planta, à atmosfera e às propriedades hidráulicas do solo.
O modelo apresentou uma boa concordância com um modelo analítico que aplica uma condição de contorno linear, à superfície da raiz, de extração de solutos dependente da concentração no solo.
A vantagem do modelo numérico sobre o analítico é que ele permite simular fluxos transientes de água e solutos, sendo, portanto, possível simular uma maior gama de situações.
Se faz necessário simulações com diferentes cenários e comparações com dados experimentais para se verificar a performance do modelo e, possivelmente, sugerir melhorias.

}

\keywordsEN {%
Solute transport; Michaelis-Menten; transient solute flux
}
\keywordsPT {%
Transporte de solutos; Michaelis-Menten; Fluxo transiente de solutos
}


%%%%% <--   % The place for your own macros is here.
\def\emph#1{{\it #1}}
\def\ttb{\tt\char`\\} %  sequence control printing of Tables
\def\asp#1{``#1''} % facilitates the use of quotation marks
\def\clim{$C_{lim}$}
\def\hlim{$H_{lim}$}
\def\c2{$C_2$}
\def\im{$I_m$}
\def\km{$K_m$}

%definition to colorize the text
\def\tred {\localcolor\Red}
\def\tblue {\localcolor\Blue}


\addprotect\ref

%\draft     % Uncomment this if the version of your document is working only.
%\linespacing=1.7  % uncomment this if you need more spaces between lines
                   % Warning: this works only when \draft is activated!
%\savetoner        % Turns off the lightBlue or lightGrey backround of
                   % verbatims, only for \draft version.
%\blackwhite       % Use this if you need really Black+White thesis.

\makefront  % Mandatory command. Makes title page, acknowledgment, contents etc.

% Files where the source of the document is prepared.
% Full name is: body.tex, appendicies.tex, the suffix can be omitted.

%\input epsf % Works onty if pdfoutput = 0
%\input body

\input introduction

\input literature

\input theory

\input methodology

\input results

\input conclusions

{\makebib{mybib}}

\input appendicies

\bye
